Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-07-03 patchy-selection-paper.tex
figure legend

The convergent evolution of similar phenotypes 
in response to shared selection pressures is a testiment to the power of selection 
to repeatedly sculpt phenotypic variation. 
In some cases this convergence extends to the molecular level, 
with disparate taxa converging to the same phenotype through parallel 
genetic changes in the same pathway, genes, 
or even by precisely the same genetic changes
\citep{stern2013genetic,zhen2012parallel,martin2013repeated}. 
Convergent adaptation can also occur within species, if individuals within a species 
adapt to the same environment, in parallel, through different genetics changes. 
There are a growing number of examples of this in a range of well studied organisms and phenotypes \cite{Arendt:08}.
Such evolution of convergent phenotypes is favored by higher mutational input, 
i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
The geographic distribution of populations can also affect the probability of parallel mutation within a species:
a widespread species is more likely to adapt by multiple, parallel mutations if dispersal is geographically limited, 
since subpopulations will adapt via new mutations before the adaptive allele arrives via migration \citep{ralph2010parallel}. 
Standing variation for a trait can also increase the probability of convergence, 
as standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
(beginning from a base of multiple copies),
leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text.
  \label{fig:mice_prob_parallel}
  }
\end{figure}

>>>>

The convergent evolution of similar phenotypes 
in response to shared selection pressures is a testiment to the power of selection 
to repeatedly sculpt phenotypic variation. 
In some cases this convergence extends to the molecular level, 
with disparate taxa converging to the same phenotype through parallel 
genetic changes in the same pathway, genes, 
or even by precisely the same genetic changes
\citep{stern2013genetic,zhen2012parallel,martin2013repeated}. 
Convergent adaptation can also occur within species, if individuals within a species 
adapt to the same environment, in parallel, through different genetics changes. 
There are a growing number of examples of this in a range of well studied organisms and phenotypes \citep{Arendt:08}.
Such evolution of convergent phenotypes is favored by higher mutational input, 
i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
The geographic distribution of populations can also affect the probability of parallel mutation within a species:
a widespread species is more likely to adapt by multiple, parallel mutations if dispersal is geographically limited, 
since subpopulations will adapt via new mutations before the adaptive allele arrives via migration \citep{ralph2010parallel}. 
Standing variation for a trait can also increase the probability of convergence, 
as standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
(beginning from a base of multiple copies),
leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics{Lava_flow_mice_prob_parallel}
  \end{center}
  \caption{
Top panel: The probability of parallel mutation as a function of the distance
between environmental patches for two different cline widths
and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
see text. Bottom panel: The initial genetic length of the haplotype
shared between patches due to adaptation via migration as a function
of the distance between environmental patches. 
  \label{fig:mice_prob_parallel}
  }
\end{figure}
